---
title: "Module 5 Report"
author: "R-Duino - Chris, Joe, Karim, Brianna"
date: "10/14/2022"
output: html_document
bibliography: S0031938497005465.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
```{css, echo=FALSE}
h1, h4 {
  text-align: center;
  color: blue;
  font-weight: bold;
}
```

## Introduction 
The central theme of this project will be to assess the impact of activity level on body temperature across various human subjects. Past literature from Chin Leong Lim shows that Gastrointestinal temperature increases by a few degrees Celsius with activity as a result of the large amount of mechanical energy converted to heat during exercise. This icnrease rarely ever exceeds 1 or 2 degrees Celsius though as a result of thermoregulation. Based on this study, we hypothesize that we will see  increases in activity correlate with small increases in body temperature in humans, but nothing larger than a few degrees Celsius as a result of the role of thermoregulatory mechanisms in maintaining physiological homeostasis. 

## Methods
To measure the body temperature of our human subjects, we  programmed an Arduino Uno unit (using the code provided in the module 5 project outline) to measure and store  temperature data to be obtained using a DHT11 sensor attached to the unit. Then, each group member had their temperature measured over 100 seconds while sitting on the couch, and then over 100 seconds after 5 minutes of jumping jacks. This data was then separately analyzed to create a table summarizing the mean temperature for each group member during each activity plus or minus 1 standard deviation. 



## Results and Discussion
```{r, echo=FALSE, message = FALSE, fig.align ='center', fig.cap= "Fig1: Figure 1 depits a box plot for body temperature data of all group members performed while at rest and after one minute of activity (sitting on couch and one minute of jumping jacks). Means are denoted by black dot."}
dat.f <- list.files(pattern = "Couch|Active.csv")

dat.l <- list() 

for(i in dat.f){
  met.dat<- unlist(strsplit(i,"_")) 
  who <- met.dat[1] 
  activity <- gsub(".csv","",met.dat[2]) 
  dat.l[[i]]<- read_csv(i)%>%
    mutate(who=who,activity=activity) %>%
    group_by(who, activity) %>% 
    mutate(mean = mean(degC))
}
dat <- do.call(rbind,dat.l) 

dat%>%
  ggplot(aes(x=activity,y=degC))+geom_boxplot()+stat_summary(fun = "mean", geom = "point", shape = 20, size = 4, fill = "white")+facet_grid(.~who) 
```

The comparison of the data allows us to determine that activity level and body temperature do seem to have an impact on one another. Based on the information presented in the box plot above, it appears that increased activity did correlate with increased body temperature, as seen with Joe, Chris, and Karim's data. Although Brianna's data does not show that activity correlated with an increase in body temperature, errors in the experimental procedure or with the way the temperature was obtained could have caused there to be no noticeable increase. Additionally, differences in physical fitness and exertion could have caused there to be no noticeable increase in body temperature for her. Based on our findings, we can conclude that body temperature and activity level exhibit a strong positive correlation with each other in humans. This finding is congruent with previous findings, as seen with the work from Chin Leong Lim.



## Contributions
Chris Peters: Assembling arduino sketch and hardware, setting up and formatting markdown code, creating the code for the plot
Joe:Found study about the effect of activity on body temperature in mice, wrote the introduction, methods,  results and discussion sections.
Karim:
Brianna



## References

